---
title: "Statistical Inference Course Project Part1"
author: "Yao Bin"
date: "June 12, 2015"
output: word_document
---
# Overview

- In this study, we investigate the exponential distribution in R and compared it with the Central Limit Theorem. The distribution of averages of 40 exponentials are investigated.  
- Code chunks of the data analysis and plots are listed in appendix section in the end of this document.

# Simulations

The exponential distribution can be simulated in R with rexp(n,lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

In this study, we set n=40 and lambda=0.2.  
Generate 1000 simulations of averages of 40 random exponentials: (see Chunk1)
```{r chunk1, echo=FALSE}
# set the parameters
n=40
simu=1000
lambda=0.2

# simulation
exp<-NULL
for(i in 1:simu) exp<-c(exp, mean(rexp(n=n, rate=lambda)))
```

# Sample Mean versus Theoretical Mean  

As mentioned above, the theoretical mean mean_t=1/lambda
```{r chunk2, echo=FALSE, comment="    "}
# theoretical mean
mean_t<-1/lambda
# sample mean
mean_s<-round(mean(exp),4)
paste("Theoretical mean mean_t=", mean_t)
paste("Sample mean mean_s=", mean_s)
```  
The sample mean **```r mean_s```** is close to theoretical mean **```r mean_t```**.

Here, we draw a plot to show that the sample mean compared with the theoretical mean.    
```{r chunk3, echo=FALSE, fig.height=4, fig.width=6}
hist(exp, prob=T, col="lightgreen",xlab="Sample Mean", 
     main="Histogram of Sample Mean versus Theoretical Mean")
# plot the sample mean
abline(v=mean_s, col="darkgreen", lwd=3)
# plot the theoretical mean
abline(v=mean_t, col="red", lty=2, lwd=3)
legend("topright", c("Sample", "Normal"), col=c("darkgreen", "red"), lwd=2)
```  

# Sample Variance versus Theoretical Variance

The theoretical variance for the exponential distribution is sigma^2 =(1/lambda)^2/n.
The sample variance is calculated by var() function in R.
```{r chunk4, echo=FALSE, comment="    "}
# theoretical variance
var_t<-(1/lambda)^2/n
# sample variance
var_s<-round(var(exp),4)
paste("Theoretical variance var_t =", var_t)
paste("Sample variance var_s=", var_s)
```    
Also, the sample variance **```r var_s```** is close to theoretical variance **```r var_t```**.

# Distribution

Here, 2 methods were performed to assess if the sample means were roughly normally distributed.(see Chunk5)  

##1. Standard normal density curve
Draw a normal density curve over the distribution of the sample means, and compare it with the distribution density curve.  
```{r chunk5.1, echo=FALSE, fig.height=4, fig.width=6}
# plot histogram of the sample means
hist(exp, prob=T, col="lightgreen",xlab="Sample Mean", main="Histogram of Sample Mean")
# plot sample distribution
lines(density(exp), lwd=2, col="darkgreen")
# plot normal distribution
curve(dnorm(x, mean=mean(exp), sd=sd(exp)), add=T, lwd=2, col="red")
legend("topright",c("Sample", "Normal"), col=c("darkgreen", "red"), lwd=2)
```

The normal density curve mostly matched the distribution density curve.  

##2. Q-Q plot  
Quantile-quantile plot (Q-Q plot) is also a widely used method to check the normality of a distribution.  
```{r chunk5.2, echo=FALSE, fig.height=4,fig.width=6}
qqnorm(exp)
qqline(exp, col="red", lwd=2)
```

The Q-Q plot shows the points scatter along the line, which indicates that this distibution is close to normal distribution.  

In conculsion, the distribution of the sample means were approximately a normal distribution.   

# Appendix: Code Chunks  
##Chunk1. Parameter setting and simulations  
```{r 1, eval=FALSE}
# set the parameters
n=40
simu=1000
lambda=0.2
# simulation
exp<-NULL
for(i in 1:simu) exp<-c(exp, mean(rexp(n=n, rate=lambda)))
```

##Chunk2. Calculating theoretical mean and sample mean  
```{r 2, eval=FALSE}
# theoretical mean
mean_t<-1/lambda
# sample mean
mean_s<-mean(exp)
mean_t;mean_s
```

##Chunk3. Plotting theoretical mean and sample mean  
```{r 3, eval=FALSE}
hist(exp, prob=T, col="lightgreen",xlab="Sample Mean", 
     main="Histogram of Sample Mean versus Theoretical Mean")
# plot the sample mean
abline(v=mean_s, col="darkgreen", lwd=3)
# plot the theoretical mean
abline(v=mean_t, col="red", lty=2, lwd=3)
legend("topright", c("Sample", "Normal"), col=c("darkgreen", "red"), lwd=2)
```

##Chunk4. Calculating theoretical variance and sample variance  
```{r 4, eval=FALSE}
# theoretical variance
var_t<-(1/lambda)^2/n
# sample variance
var_s<-var(exp)
var_t;var_s
```

##Chunk5. Distribution test  
```{r 5, eval=FALSE}
# plot histogram of the sample means
hist(exp, prob=T, col="lightgreen",xlab="Sample Mean", main="Histogram of Sample Mean")
# plot sample distribution
lines(density(exp), lwd=2, col="darkgreen")
# plot normal distribution
curve(dnorm(x, mean=mean(exp), sd=sd(exp)), add=T, lwd=2, col="red")
legend("topright",c("Sample", "Normal"), col=c("darkgreen", "red"), lwd=2)

# Q-Q plot
qqnorm(exp)
qqline(exp, col="red", lwd=2)
```


